---
title: "penman_montieth_jarvis"
author: "Bar Avni"
date: "11/14/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Penman and Montieth and Modification of Jarvis equation model

###This code calls the Jarvis_mod function, generate the gs variable and then call the penman_montieth function to generate ET. In the next stage, sensetivity analysis are perform.

```{r}
# 1. Load data:
Tdata = read.table("metdata.txt",header = TRUE, sep = "")

# Calculate the maximum surface conductance:
gsmax = max(Tdata[,c("gs")])

# Call the T variable from the metata file:
Tair = Tdata[,c("tavg")]
```

###The modified Jarvis function:

``` {r}
#'  Modified Jarvis Equation
#'  
#'  This funciton compute surface conductance based on the Jarvis equation, 
#'  as function of tempature.


#'  @param Tair (deg C) leaf temperature
#'  @param Kt 
#'  @param Tx 
#'  @param gsmax (s/mm)  maximum surface conductance - the maximum stomatal conductance of the plant

#'  @author Bar Avni
#'  @return gs surface conductance (s/mm)

Jarvis_mod =
  function(Tair,gsmax, Kt, Tx) {
    
    #       Internal Variables
    
    #       tk       (deg K)   
    #       Ta       (deg K) 

    # A simple representation of  tempature function in Jarvis equation (Dickinson, 1984)
    tk = Tair + 273.15 
   

    Ta = (1-Kt*(Tx-tk)^2)^(-1)
    
    for (i in 1:length(Ta)) {
    if (Ta[i] > 0) {
      Ta[i] = 1 
     } else {
      Ta[i] = 0 
     }
    }
   
    #Ta = Ta + 273.15
    gs = gsmax*Ta
    # return from your function
    return(gs)
  }
``` 

Call the jarvis_modfunction:
``` {r}
gs <- Jarvis_mod(Tair,gsmax = 1.710054, Kt = 0.0016, Tx = 298)
```
I used only the tempature variable from modified Jarvis equation. The relation of surface conductance to tempature is taken from Lhomme (1998).
$f(T) = [1-Kt(Tx-T)^{2}]^{-1}$  
where Kt is 0.0016, Tx is 298 K and T is air tempature.
The temapture changes throughout the year according to the seasons and climate and therefore the surface conductance changes. I simplified the conductance behivior into two states: on or off depends on the temapture.

```{r}
# load the variables for Penman&Montieth function: 
vpd = Tdata[,c("vpd")] 
Rnet = Tdata[,c("rnet")]
ga = Tdata[,c("ga")]
dayl = Tdata[,c("ga")]
```

###The Penman&Montieth function:

```{r}
#' Penman Montieth Equation
#' 
#' THis function computer evapotranspiration based on radiation, conductance etc
#' 

#' @param     Tair    (deg C) air temperature
#' @param     vpd     (Pa)    vapour pressure deficit
#' @param     Rnet    (j/m2/day)      net radiation
#' @param     gs      (s/mm)  surface conductance
#' @param     ga      (s/mm)  aerodynamic conductance
#' @param     CP  =      1010.0 (J/kg*K) specific heat of air
#' @param     Pair = 101325 (Pa)
#' @author Naomi
#' @return Evapotranspiration (mm/day)


penman_montieth =
  function(Tair, vpd, Rnet, gs,ga, dayl, CP=1010, Pair=101325) {
    
    #       Internal Variables
    #
    #       rho     (kg/m3)         density of air
    #       CP      (K/jg/degC)     specific heat of air
    #       lhvap   (J/kg)          latent heat of vapourization H20
    #       s       (Pa/degC)       slope of sat vpd vs T curve
    #       rs      (s/m)   surface resistance
    #       ra      (s/m)   aerodynamic resistance
    
    
    # convert Rnet to daytime value in j/m2/s
    Rnet = Rnet / (60*60*dayl)
    
    # convert conductance to resistance and change units
    rs = 1000.0/gs
    ra = 1000.0/ga
    
    # Assign tk (Kelvins)
    tk = Tair + 273.15
    
    #       Density of air (rho) as a fn. of air temp.
    rho = 1.292 * ( 0.00428 * Tair )
    
    #       Latent heat of vapourization as a fn. of Tair.
    lhvap = 2.5023e6 - 2430.54 * Tair
    
    #       Temperature offsets for slope estimates
    dt = 0.2
    t1 = Tair + dt
    t2 = Tair - dt
    
    #       Saturation vapour pressures at t1 and t2(Pa)
    pvs1 = 610.7 * exp(17.38 * t1 / ( 239.0 + t1))
    pvs2 = 610.7 * exp(17.38 * t2 / ( 239.0 + t2))
    
    #       Slope of pvs vs T curve at Tair (Pa/deg C)
    s = ( pvs1 -   pvs2 ) / ( t1 -  t2 )
    
    #       Calculate gamma
    gamma = CP * Pair / ( lhvap )
    
    
    #       Evaporation in W/m2
    et = ((s*Rnet) + (rho*CP*vpd/ra)) / (gamma*(1.0 + rs/ra) +s)
    
    
    #       mH20/s = W/m2 * 1kgH20/lhvap J * 1m3H20/1000kGH20
    ewater = ( et/ ( lhvap * 1000 ))
    
    # mmH20/day
    ewater.day = ewater * dayl*60*60 * 1000
    
    # return from your function
    ewater.day
  }

```

Call the Penman&Montieth function

```{r}
Et = penman_montieth(Tair, vpd, Rnet, gs,ga, dayl = 24, CP=1010, Pair=101325)

# Perform sensitivity analysis to eximine how increasing air tempature impacts ET estimates
# by multiply the air tempature by factor of -20:
Tfactor = -20
for (i in 1:length(Tfactor)) {
  Tair_fac2 = Tair*Tfactor[i]
  gs_fac2 <- Jarvis_mod(Tair_fac2, gsmax = 1.710054, Kt = 0.0016, Tx = 298)
}

```
# Plot ET valuse as function of time:

```{r, echo=FALSE}
#plot graphs:
date = Tdata[,c("date")]
plot(date,Et,type = "l", col = "red", xlab = "Time", ylab = "Evapotranspiration (mm/day)")
# Create a title with a red, bold/italic font
title(main="ET valuse as function of time", col.main="black", font.main=4)
```

# Add figures to visualize the two scenarios:

```{r, echo=FALSE}
#plot graphs:
plot(Tair,gs,type = "l", col = "red", xlab = "air tempature (K)", ylab = "surface conductance (s/mm)")
# Create a title with a red, bold/italic font
title(main="gs as function of air tempature", col.main="black", font.main=4)
plot(Tair_fac2,gs_fac2,type = "l", col = "blue", xlab = "air tempature (K)", ylab = "surface conductance (s/mm)")
title(main="gs as function of air tempature times factor -20", col.main="black", font.main=4)
```


The model is not sensitive to temapture due to the assumption was done in the Jarvis_mod function. Since the surface conductance have not change after implimanting the factor, there is no need to run the penman&montieth model.

