---
title: "Thornthwaite"
author: "Bar Avni"
date: "11/15/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
## Research question:  
 
 How the monthly potential ET is affected by the tempature within and between months. 
 
## methods:   
 
The Thornthwaite model was developed by Warren Thornthwaite in 1948 to estimate the potential evapotranspiration. Potential evaporation or potential evapotranspiration (PET) is defined as the amount of evaporation that would occur if a sufficient water source were available. If the actual evapotranspiration is considered the net result of atmospheric demand for moisture from a surface and the ability of the surface to supply moisture, then PET is a measure of the demand side. Surface and air temperatures, insolation, and wind all affect this. The Thornthwaite model using the mean tempature as measured for each month, to calculate the monthly heat index. The annual heat index is calculating next by summing up the monthly heat index for each year and multiply it by factor. Lastly, thr monthly potential evapotranspiration is calculating following the equation:  
$Ep = 1.6*b(10t/I)^{a}$ 
Where I is the annual heat index, b is the total monthly daylight hours divided by 360, a is a cubic function of I and t is the mean temperature of the month (dec C).  
 


### Load data:
```{r }
Tdata = read.table("metdata.txt",header = TRUE, sep = "")
day = Tdata[,c("day")]
Tair = Tdata[,c("tavg")]
year = Tdata[,c("year")]
data_tmdh = read.table("TMDH.txt",header = TRUE, sep = "")
TMDH = data_tmdh[,c("TMDH")]
month_tmdh = data_tmdh[,c("month")]
```

### The Thornthwaite function is written in R:
 
```{r }

# Thornthwaite Model
# Developed by Warren Thornthwaite in 1948 to estimate PET:

#' @param     day
#' @param     year    
#' @param     Tair    (deg C) air temperature
#' @param     TMDH    Total monthly daylight hours

#' @author Bar Avni 
#' @return monthly potential ET 

Thornthwaite = 
  function(day, year, Tair, TMDH){
    
    #       Internal Variables
    
    #       day_diff = the diffrences between months
    #       day_diff0 = find the first day of the month
    #       Tmonth = Mean temperature of the month (deg C)
    #       MH_index = Monthly heat index (deg C)
    #       year_diff = find the first day of the year
    #       year_diff1 = find the first day of the year
    #       year_intersect = thake the index of the first day of the year
    #       AH_factor = multiple the month index to create the annual index
    #       AH_index = annual index
    #       a = Thornthwaite parameter
    #       b = Thornthwaite parameter
    #       Ep = monthly potential ET 
    

# Devide the manths in order to calculate the mean monthly tempature:
day_diff = diff(day)
for (i in 1:length(day_diff)) {
  if (day_diff[i] < 1) {
    day_diff[i] = 0 
  }
}
  
day_diff0 = which(day_diff %in% 0)

for (j in 1:length(day_diff0)) {
  if (j == 1){
    Tmonth = mean(Tair[1:day_diff0[j]])  
  } else {
    Tmonth[j] = mean(Tair[(day_diff0[j-1]+1):day_diff0[j]])  
  }
}
# monthly heat index:
MH_index = (Tmonth/5)^(1.514)

# Sum the 12 monthly heat index into an annual heat index (I):
year_diff = diff(year)

year_diff1 = which(year_diff %in% 1)
year_intersect = intersect(year_diff1, day_diff0)

year_index = which(day_diff0 %in% year_intersect)

AH_factor = (MH_index/5)^(1.514)

for (n in 1:length(year_index)) {
  if (n == 1){
    AH_index = sum(AH_factor[1:year_index[n]]) 
  } else {
    AH_index[n] = sum(AH_factor[(year_index[n-1]+1):year_index[n]])  
  }
}

# Monthly potential ET - Ep:
Ep = 0
for (m in 1:length(AH_index)) {
  a = 0.49+1.79*10^(-2)*AH_index[m] + 7.711*10^(-5)*AH_index[m]^2 + 6.751*10^(-7)*AH_index[m]^3
  for(l in 1:length(MH_index)){
    b =  TMDH[l]/360
    Ep[l] = 1.6*b*((10*Tmonth[l])/(AH_index[m]))^a
  }
}

Ep
}

    
```
 
## Applicaiton of the function: 
I used the metdata.txt file from the course gitHub. The function takes dates, air temperature,  and monthly daylight hours as inputs and return the monthly ET potential.  

### Call the function:
``` {r}
Ep = Thornthwaite(day, year, Tair, TMDH)
```

## Results:
The Evapotranspiration monthly potential is plotted for years 2001, 2003 and 2008: 

```{r pressure, echo=FALSE}
x = month_tmdh[4:15]
y1 = Ep[4:15]
y2 = Ep[28:39]
y3 = Ep[88:99]

plot(x, y1, type = "p", col = "red", xlab = "month", ylab = "Ep (cm)")
par(new = TRUE)
plot(x, y2, type = "p", col = "green")
par(new = TRUE)
plot(x, y3, type = "p", col = "blue")
# Create a title with a red, bold/italic font
title(main="Monthly potential ET for 2001, 2003, 2008", col.main="black", font.main=4)
legend("topleft", legend=c("2001", "2003", "2008"),
       col=c("red", "green", "blue"), pch = 1, cex=0.5)
```

From these results we can cerfuly say that the Ep of 2008 was the highets and 2003 had the lowes Ep values. May and July in 2001 had higher Ep values than the those months in 2008. Augost had the highest Ep values overall. 